---
title: "US State Vaccinations"
author: "Jefferson Santos - ANA 515"
date: '11/09/2021'
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
```

This dataset provides information about all the vaccinations by US State. 
It has been collected from two main websites: https://ourworldindata.org/us-states-vaccinations and https://covid.ourworldindata.org/data/vaccinations/us_state_vaccinations.csv
It has vaccines data like: total distributed, people vaccinated, daily vaccinations e share doses used.This is saved as .csv file (comma-separated file). We can open it using microsoft excel or any text editor.

```{r}
# using read.csv function and assigning it to an object
# We will use the built in read.csv(...) function call, which reads the data in as a data frame, and assign the data frame to a variable (using <-) so that it is stored in R’s memory.
data <- read.csv('us_state_vaccinations.csv')
```

```{r results = TRUE}
# getting the column names
colnames(data)
```

```{r results = FALSE}
# renaming columns
data <- rename(data, state = location)
```

```{r results = FALSE}
# removing few columns from the dataset
data_subset <- select(data, date,	state,	total_vaccinations, total_distributed, people_fully_vaccinated, daily_vaccinations, share_doses_used)
```

```{r results = TRUE}
# getting the column names again to confirm the changes
colnames(data_subset)
```

```{r results = TRUE}
# getting the dimensions of the data subset
dimensions <- dim(data_subset)
```

This dataframe has `r dimensions[1]` rows and `r dimensions[2]` columns. The names of the columns and a brief description of each are in the table below:

```{r results = TRUE}
library(knitr)

columns_summary <- data.frame(
  Columns = c(colnames(data_subset)),
  Description = c(
    "Date of the information been provided",
    "US State/Location",
    "Cumulative vaccinations at that date per state",
    "Cumulative distribuition at that date per state",
    "Cumulative people fully vaccinated at that date per state",
    "Total of people vaccinated on that date per state",
    "The ratio between Cumulative distribuition by Cumulative vaccinations"
  )
)

kable(columns_summary, caption = "US State Vaccinations Columns Description")
```

```{r results = FALSE}
#filtering the data by location = entire US instead of by State to avoid an incorret summary below
data_us <- filter(data_subset, state == "United States")

# picking three columns to use summary function
data_pick3 <- select(data_us, people_fully_vaccinated, daily_vaccinations, share_doses_used)

# Summary functions (min, max, mean, missing values)
pfv_min <- format(big.mark=",", scientific=FALSE,round(min(data_pick3$people_fully_vaccinated, na.rm=TRUE),2), nsmall = 2)
pfv_max <- format(big.mark=",", scientific=FALSE,round(max(data_pick3$people_fully_vaccinated, na.rm=TRUE),2), nsmall = 2)
pfv_mean <- format(big.mark=",", scientific=FALSE,round(mean(data_pick3$people_fully_vaccinated, na.rm=TRUE),2), nsmall = 2)
pfv_sumna <- sum(is.na(data_pick3$people_fully_vaccinated))

dv_min <- format(big.mark=",", scientific=FALSE,round(min(data_pick3$daily_vaccinations, na.rm=TRUE),2), nsmall = 2)
dv_max <- format(big.mark=",", scientific=FALSE,round(max(data_pick3$daily_vaccinations, na.rm=TRUE),2), nsmall = 2)
dv_mean <- format(big.mark=",", scientific=FALSE,round(mean(data_pick3$daily_vaccinations, na.rm=TRUE),2), nsmall = 2)
dv_sumna <- sum(is.na(data_pick3$daily_vaccinations))

sdu_min <- format(big.mark=",", scientific=FALSE,round(min(data_pick3$share_doses_used, na.rm=TRUE),2), nsmall = 2)
sdu_max <- format(big.mark=",", scientific=FALSE,round(max(data_pick3$share_doses_used, na.rm=TRUE),2), nsmall = 2)
sdu_mean <- format(big.mark=",", scientific=FALSE,round(mean(data_pick3$share_doses_used, na.rm=TRUE),2), nsmall = 2)
sdu_sumna <- sum(is.na(data_pick3$share_doses_used))
```

```{r results = TRUE}
summary_stats <- data.frame(
  Summary = c(colnames(data_pick3)),
  Min = c(pfv_min,
          dv_min,
          sdu_min
  ),
  Max = c(pfv_max,
          dv_max,
          sdu_max
  ),
  Mean = c(pfv_mean,
          dv_mean,
          sdu_mean
  ),
  NAs = c(pfv_sumna,
          dv_sumna,
          sdu_sumna
  )
)

kable(t(summary_stats), caption = "US Vaccinations Summary Stats")
```
